---
output:
  word_document: default
  html_document: default
  pdf_document: default
---
# Intro to Dat Science - HW 2
##### Copyright  Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Rutwik Ghag
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

### Reminders of things to practice from last week: 
Assignment arrow  		<- <br>
The combine command 	c( ) <br>
Descriptive statistics 	mean( ) sum( ) max( ) <br>
Arithmetic operators		+ - * /<br>
Boolean operators		>    <    >=    <=   ==   !=  <br>

**This Week:** Explore the **quakes** dataset (which is included in R). Copy the **quakes** dataset into a new dataframe (call it **myQuakes**), so that if you need to start over, you can do so easily (by copying quakes into myQuakes again). Summarize the variables in **myQuakes**. Also explore the structure of the dataframe

library(tidyverse)

```{r}
myQuakes <- quakes
summary(myQuakes)
str(myQuakes)
```

**Step 1:** Explore the earthquake magnitude variable called **mag**  <br>

A.	What is the average magnitude? Use mean() or summary():


```{r}
mean(myQuakes$mag)
```

B.	What is the magnitude of the largest earthquake? Use max() or summary() and save the result in a variable called **maxQuake**:


```{r}
maxQuake <- max(myQuakes$mag)
maxQuake
```

C.	What is the magnitude of the smallest earthquake? Use min() or summary() and save the result in a variable called **minQuake**:


```{r}
minQuake <- min(myQuakes$mag)
minQuake
```

D.	Output the **third row** of the dataframe


```{r}
myQuakes[3,]
```

E. Create a new dataframe, with only the rows where the **magnitude is greater than 4**. How many rows are in that dataframe (use code, do not count by looking at the output) 


```{r}
newQuakes <- myQuakes[myQuakes$mag>4,]
nrow(newQuakes)
```

F.	Create a **sorted dataframe** based on magnitude and store it in **quakeSorted1**. Do the sort two different ways, once with arrange() and then with order()


```{r}
#quakeSorted1 <- myQuakes %>% arrange(mag)
#quakeSorted1

# Added to comments because it was not recognizing the %>% symbols and was unable to knit

quakeSorted1 <- myQuakes[order(myQuakes$mag),]
quakeSorted1
```

G.	What are the latitude and longitude of the quake reported by the largest number of stations?  


```{r}
latlon<-myQuakes[which.max(myQuakes$stations),]
latlon[,1:2]
```

H.	What are the latitude and longitude of the quake reported by the smallest number of stations?  


```{r}
latlon2 <- myQuakes[which.min(myQuakes$stations),]
latlon2[,1:2]
```

**Step 3:** Using conditional if statements <br>

I.	 Test if **maxQuake** is greater than 7 (output “yes” or “no”) <br>
**Hint:** Try modifying the following code in R: 


```{r}
if  (maxQuake>7) 
  print("Yes" ) else 
  print("No")

```

J.	 Following the same logic, test if **minQuake** is less than 3 (output “yes” or “no”):


```{r}
if (minQuake < 3)
  print("Yes") else
    print("No")
```
