---
output:
  word_document: default
  html_document: default
---
# Intro to Data Science - HW 5

##### Copyright 2021, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Rutwik Ghag
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).


```{r}
library(RCurl) # install the RCurl library
library(jsonlite) # install the JSON library
dataset <- getURL("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json") # copy the contents of the JSON file from the mentioned URL into the variable
readlines <- jsonlite::fromJSON(dataset) # convert the JSON data into an R object 
df <- readlines$objects$person # store the data in readlines in a data frame
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).


```{r}
head(df)
str(df)
#View(df)
```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>



```{r}
# The dataset is about senators and their information like name, gender, twitter id
```

## Part 2: Investigate the resulting dataframe 

C.	How many senators are women? 


```{r}
sum(df$gender=="female")

```

D.	How many senators have a YouTube account?  


```{r}
sum(!is.na(df$youtubeid))
```

E.	How many women senators have a YouTube account?


```{r}
sum(df$gender=="female" & !is.na(df$youtubeid))
```

F.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   


```{r}
youtubeWomen <- df[df$gender=="female" & !is.na(df$youtubeid),]
```

G.	What does running this line of code do? Explain in a comment: 


```{r}
youtubeWomen$year <- substr(youtubeWomen$birthday,1,4)
# adds a new column, year, in the df youtubeWomen to store the birth year
```

H.	Use this new variable to calculate the mean **birthyear** in **youtubeWomen**. <br>
**Hint:** You may need to convert it to numeric first.


```{r}
mean(as.numeric(youtubeWomen$year))

```

I.	Make a histogram of the **birthyears** of senators in **youtubeWomen**. Add a comment describing the shape of the distribution.


```{r}
hist(as.numeric(youtubeWomen$year))
#the histogram is like an evenly distributed bell curve 
```
